---
title: "Finance Econometrics"
author: "André Pellizzaro"
date: "`r Sys.Date()`"
output: html_notebook
---
# Course Outline
The objective of the course is to present linear and non-linear models for both conditional mean and conditional variances. The Extreme Values Theory, Theory of Copulas and Additional Topics (Models for High Frequency Data, Continuous Time Modelsor High Dimensional Model) will also be presented.

## Stylized Facts in Financial Data

Primeiro, iremos buscar as bases de dados, sendo `c("^IXIC","AAPL")` o índice NASDAQ e a ação da Apple, respectivamente:

```{r message=FALSE, warning=FALSE}
rm(list=ls())
library(tidyquant)
library(tidyverse)
df <- tq_get(c("^IXIC","AAPL"), get = "stock.prices", from = "2016-01-01", complete_cases = FALSE)
```

Faremos então o **retorno composto** e para o fechamento do dia (Close):

```{r message=FALSE, warning=FALSE}
ts <- df |> 
  group_by(symbol) |> 
  summarise(date = date, 
            close = close , 
            lClose = log(close), 
            rsClose = ((close-lag(close))/lag(close))*100,
            rcClose = (lClose - lag(lClose))*100
              )
ts |> group_by(symbol) |>
    ggplot(aes(x = date, y = rcClose, color = symbol)) +
    geom_line(size = 1) +
    labs(title = "Retorno Composto",
         x = "", y = "Retorno", color = "") +
    facet_wrap(~ symbol, ncol = 2, scales = "free_y") +
#    scale_y_continuous(labels = scales::dollar) +
    theme_tq() 
```




Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
